---
title: "Cost of functional independence limitations"
author: "RSF"
date: "2022-07-28"
output: 
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1) Set-Up
Data management for setting up the cost tables
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(openxlsx, dplyr, robumeta,metafor, clubSandwich, ggplot2, psych, naniar)

setwd("~/Desktop/Manuscripts/Ongoing Projects/SR on Economic Impact of Frailty")
data1<-read.xlsx("Cost data for meta FINAL.xlsx")


data1<-rename(data1, c("Study_Length"="Length.(Years)"))

data1$Annual.Costs.mean <- data1$Costs.mean/data1$Study_Length
data1$Annual.Costs.sd <- data1$Costs.SD/data1$Study_Length

data1$Cost.Perspective<-NA
data1$Cost.Perspective[data1$Cost.persective == "Social" | data1$Cost.persective == "Societal" | data1$Cost.persective == "Societal "] <- "Societal"
data1$Cost.Perspective[data1$Cost.persective == "Healthcare"] <- "Healthcare"
data1$Cost.Perspective[data1$Cost.persective == "Personal" | data1$Cost.persective == "Patient"] <- "Personal"

data2<- subset(data1, !is.na(Cost.Perspective))

data3<- data2 %>%
  mutate(Annual.Costs.sd.liberal = ifelse(is.na(Annual.Costs.sd), Annual.Costs.mean, Annual.Costs.sd),
         Annual.Costs.sd.conservative = ifelse(is.na(Annual.Costs.sd), Annual.Costs.mean*2, Annual.Costs.sd))

data3$costs.subtype.2<-NA
data3$costs.subtype.2[data3$Cost.Subtype == "Clinical" | data3$Cost.Subtype == "Clinician"]<-"Clinician"
data3$costs.subtype.2[data3$Cost.Subtype == "Hospital costs" | data3$Cost.Subtype == "Hospital"]<-"Hospital"
data3$costs.subtype.2[data3$Cost.Subtype == "All" | data3$Cost.Subtype == "Total"]<-"All"
data3$costs.subtype.2[data3$Cost.Subtype == "Assisted Services" | data3$Cost.Subtype == "Day care"]<-"Assisted Services"
data3$costs.subtype.2[data3$Cost.Subtype == "Direct Health Care Related Costs"]<-"Direct Healthcare Related Costs"
data3$costs.subtype.2[data3$Cost.Subtype == "Home-Help Indirect"]<-"Home-Help Indirect"
data3$costs.subtype.2[data3$Cost.Subtype == "Home Healthcare"]<-"Home Healthcare"
data3$costs.subtype.2[data3$Cost.Subtype == "Medications"]<-"Medications"
data3$costs.subtype.2[data3$Cost.Subtype == "Outpatient"]<-"Outpatient"

data3$Cost.Perspective[data3$costs.subtype.2 == "Medications" & data3$Cost.Perspective == "Healthcare"]<-"Personal"
data3$Cost.Perspective[data3$costs.subtype.2 == "Outpatient" & data3$Cost.Perspective == "Societal"]<-"Healthcare"
data3$Cost.Perspective[data3$costs.subtype.2 == "Home Healthcare" & data3$Cost.Perspective == "Personal"]<-"Societal"

data4<-data3 %>%
  mutate(Healthcare_All = ifelse(Cost.Perspective == "Healthcare" & costs.subtype.2 == "All", Annual.Costs.mean, NA),
         Societal_All = ifelse(Cost.Perspective == "Societal" & costs.subtype.2 == "All", Annual.Costs.mean, NA),
         Personal_All = ifelse(Cost.Perspective == "Personal" & costs.subtype.2 == "All", Annual.Costs.mean, NA),
         Healthcare_Clinician = ifelse(Cost.Perspective == "Healthcare" & costs.subtype.2 == "Clinician", Annual.Costs.mean, NA),
         Healthcare_Hospital = ifelse(Cost.Perspective == "Healthcare" & costs.subtype.2 == "Hospital", Annual.Costs.mean, NA),
         Healthcare_Outpatient = ifelse(Cost.Perspective == "Healthcare" & costs.subtype.2 == "Outpatient", Annual.Costs.mean, NA),
         Societal_Assisted = ifelse(Cost.Perspective == "Societal" & costs.subtype.2 == "Assisted Services", Annual.Costs.mean, NA),
         Societal_HomeHealth = ifelse(Cost.Perspective == "Societal" & costs.subtype.2 == "Home Healthcare", Annual.Costs.mean, NA),
         Personal_HealthCosts= ifelse(Cost.Perspective == "Personal" & costs.subtype.2 == "Direct Healthcare Related Costs", Annual.Costs.mean, NA),
         Personal_HomeHelp = ifelse(Cost.Perspective == "Personal" & costs.subtype.2 == "Home-Help Indirect", Annual.Costs.mean, NA),
         Personal_Meds = ifelse(Cost.Perspective == "Personal" & costs.subtype.2 == "Medications", Annual.Costs.mean, NA),
         Healthcare_All_lib_SD = ifelse(Cost.Perspective == "Healthcare" & costs.subtype.2 == "All", Annual.Costs.sd.liberal, NA),
         Societal_All_lib_SD = ifelse(Cost.Perspective == "Societal" & costs.subtype.2 == "All", Annual.Costs.sd.liberal, NA),
         Personal_All_lib_SD = ifelse(Cost.Perspective == "Personal" & costs.subtype.2 == "All", Annual.Costs.sd.liberal, NA),
         Healthcare_Clinician_lib_SD = ifelse(Cost.Perspective == "Healthcare" & costs.subtype.2 == "Clinician", Annual.Costs.sd.liberal, NA),
         Healthcare_Hospital_lib_SD = ifelse(Cost.Perspective == "Healthcare" & costs.subtype.2 == "Hospital", Annual.Costs.sd.liberal, NA),
         Healthcare_Outpatient_lib_SD = ifelse(Cost.Perspective == "Healthcare" & costs.subtype.2 == "Outpatient", Annual.Costs.sd.liberal, NA),
         Societal_Assisted_lib_SD = ifelse(Cost.Perspective == "Societal" & costs.subtype.2 == "Assisted Services", Annual.Costs.sd.liberal, NA),
         Societal_HomeHealth_lib_SD = ifelse(Cost.Perspective == "Societal" & costs.subtype.2 == "Home Healthcare", Annual.Costs.sd.liberal, NA),
         Personal_HealthCosts_lib_SD = ifelse(Cost.Perspective == "Personal" & costs.subtype.2 == "Direct Healthcare Related Costs", Annual.Costs.sd.liberal, NA),
         Personal_HomeHelp_lib_SD = ifelse(Cost.Perspective == "Personal" & costs.subtype.2 == "Home-Help Indirect", Annual.Costs.sd.liberal, NA),
         Personal_Meds_lib_SD = ifelse(Cost.Perspective == "Personal" & costs.subtype.2 == "Medications", Annual.Costs.sd.liberal, NA),
         Healthcare_All_con_SD = ifelse(Cost.Perspective == "Healthcare" & costs.subtype.2 == "All", Annual.Costs.sd.conservative, NA),
         Societal_All_con_SD = ifelse(Cost.Perspective == "Societal" & costs.subtype.2 == "All", Annual.Costs.sd.conservative, NA),
         Personal_All_con_SD = ifelse(Cost.Perspective == "Personal" & costs.subtype.2 == "All", Annual.Costs.sd.conservative, NA),
         Healthcare_Clinician_con_SD = ifelse(Cost.Perspective == "Healthcare" & costs.subtype.2 == "Clinician", Annual.Costs.sd.conservative, NA),
         Healthcare_Hospital_con_SD = ifelse(Cost.Perspective == "Healthcare" & costs.subtype.2 == "Hospital", Annual.Costs.sd.conservative, NA),
         Healthcare_Outpatient_con_SD = ifelse(Cost.Perspective == "Healthcare" & costs.subtype.2 == "Outpatient", Annual.Costs.sd.conservative, NA),
         Societal_Assisted_con_SD = ifelse(Cost.Perspective == "Societal" & costs.subtype.2 == "Assisted Services", Annual.Costs.sd.conservative, NA),
         Societal_HomeHealth_con_SD = ifelse(Cost.Perspective == "Societal" & costs.subtype.2 == "Home Healthcare", Annual.Costs.sd.conservative, NA),
         Personal_HealthCosts_con_SD = ifelse(Cost.Perspective == "Personal" & costs.subtype.2 == "Direct Healthcare Related Costs", Annual.Costs.sd.conservative, NA),
         Personal_HomeHelp_con_SD = ifelse(Cost.Perspective == "Personal" & costs.subtype.2 == "Home-Help Indirect", Annual.Costs.sd.conservative, NA),
         Personal_Meds_con_SD = ifelse(Cost.Perspective == "Personal" & costs.subtype.2 == "Medications", Annual.Costs.sd.conservative, NA),)

data5<-transform(data4, Healthcare_All = as.numeric(Healthcare_All),
         Societal_All = as.numeric(Societal_All),
         Personal_All = as.numeric(Personal_All),
         Healthcare_Clinician = as.numeric(Healthcare_Clinician),
         Healthcare_Hospital = as.numeric(Healthcare_Hospital),
         Healthcare_Outpatient = as.numeric(Healthcare_Outpatient),
         Societal_Assisted = as.numeric(Societal_Assisted),
         Societal_HomeHealth = as.numeric(Societal_HomeHealth),
         Personal_HealthCosts = as.numeric(Personal_HealthCosts),
         Personal_HomeHelp = as.numeric(Personal_HomeHelp),
         Personal_Meds = as.numeric(Personal_Meds),
         Healthcare_All_lib_SD = as.numeric(Healthcare_All_lib_SD),
         Societal_All_lib_SD = as.numeric(Societal_All_lib_SD),
         Personal_All_lib_SD = as.numeric(Personal_All_lib_SD),
         Healthcare_Clinician_lib_SD = as.numeric(Healthcare_Clinician_lib_SD),
         Healthcare_Hospital_lib_SD = as.numeric(Healthcare_Hospital_lib_SD),
         Healthcare_Outpatient_lib_SD = as.numeric(Healthcare_Outpatient_lib_SD),
         Societal_Assisted_lib_SD = as.numeric(Societal_Assisted_lib_SD),
         Societal_HomeHealth_lib_SD = as.numeric(Societal_HomeHealth_lib_SD),
         Personal_HealthCosts_lib_SD = as.numeric(Personal_HealthCosts_lib_SD),
         Personal_HomeHelp_lib_SD = as.numeric(Personal_HomeHelp_lib_SD),
         Personal_Meds_lib_SD = as.numeric(Personal_Meds_lib_SD),
         Healthcare_All_con_SD = as.numeric(Healthcare_All_con_SD),
         Societal_All_con_SD = as.numeric(Societal_All_con_SD),
         Personal_All_con_SD = as.numeric(Personal_All_con_SD),
         Healthcare_Clinician_con_SD = as.numeric(Healthcare_Clinician_con_SD),
         Healthcare_Hospital_con_SD = as.numeric(Healthcare_Hospital_con_SD),
         Healthcare_Outpatient_con_SD = as.numeric(Healthcare_Outpatient_con_SD),
         Societal_Assisted_con_SD = as.numeric(Societal_Assisted_con_SD),
         Societal_HomeHealth_con_SD = as.numeric(Societal_HomeHealth_con_SD),
         Personal_HealthCosts_con_SD = as.numeric(Personal_HealthCosts_con_SD),
         Personal_HomeHelp_con_SD = as.numeric(Personal_HomeHelp_con_SD),
         Personal_Meds_con_SD = as.numeric(Personal_Meds_con_SD))

data5$Study_Design<-NA
data5$Study_Design[data5$Study.Design == "Observational"]<-"Observational"
data5$Study_Design[data5$Study.Design == "RCT" | data5$Study.Design == "RCT "] <- "RCT"
data5$Study_Design[data5$Study.Design == "Computer Simulation" |data5$Study.Design == "Cost-effectiveness study" | data5$Study.Design == "Quasi-Experimental"] <- "Other"
```


# 2) Function for setting up cost-analyses
This step provides the means, number of studies, and 95% CIs for each cost-perspective
```{r}
CI.intvl <- function(x){
  avg <- as.numeric(mean(x, na.rm=TRUE))
  std.dev<- as.numeric(sd(x, na.rm = TRUE))
  number<- as.numeric(length(x[!is.na(x)]))
  std.err<- as.numeric(std.dev/sqrt(number))
  UL.CI<- avg + 1.96*std.err
  LL.CI<- avg - 1.96*std.err
  return(list(study_number = number, Mean=avg, Upper_Limit=UL.CI, Lower_Limit=LL.CI))
}
```

# 3) Indexing studies by cost perspectives
This step provides the different number of studies for each perspective
```{r}
Cost.perspective.counts<-data5 %>%
 count(Cost.Perspective,Study)

Cost.perspective.counts %>%
  count(Cost.Perspective)
```

This determines the number of studies which used all 3 perspectives
```{r}
Cost.perspective.count2<-Cost.perspective.counts %>%
  count(Study,Cost.Perspective)

Cost.perspective.count3<-count(Cost.perspective.count2,Study)
subset(Cost.perspective.count3,n==3)
count(subset(Cost.perspective.count3,n==3))
```


# 4) Costs for each perspective

## 4.1) Social
```{r}
#Societal Means
CI.intvl(data5$Societal_All)
CI.intvl(data5$Societal_HomeHealth)
CI.intvl(data5$Societal_Assisted)

#Societal SD Liberal 
CI.intvl(data5$Societal_All_lib_SD)
CI.intvl(data5$Societal_HomeHealth_lib_SD)
CI.intvl(data5$Societal_Assisted_lib_SD)
```

## 4.2) Healthcare
```{r}
#Healthcare Means
CI.intvl(data5$Healthcare_All)
CI.intvl(data5$Healthcare_Clinician)
CI.intvl(data5$Healthcare_Hospital)
CI.intvl(data5$Healthcare_Outpatient)

#Healthcare SD Liberal
CI.intvl(data5$Healthcare_All_lib_SD)
CI.intvl(data5$Healthcare_Clinician_lib_SD)
CI.intvl(data5$Healthcare_Hospital_lib_SD)
CI.intvl(data5$Healthcare_Outpatient_lib_SD)
```

## 4.3) Personal
```{r}
#Personal Means
CI.intvl(data5$Personal_All)
CI.intvl(data5$Personal_HealthCosts)
CI.intvl(data5$Personal_HomeHelp)
CI.intvl(data5$Personal_Meds)

#Personal SD Liberal
CI.intvl(data5$Personal_All_lib_SD)
CI.intvl(data5$Personal_HealthCosts_lib_SD)
CI.intvl(data5$Personal_HomeHelp_lib_SD)
CI.intvl(data5$Personal_Meds_lib_SD)
```


# 5) Costs based on studies which have indices of frailty, mobility, or cognition

First we stratify by the index of interest
```{r}
Frailty<-subset(data5, Frailty.Indexed == 1)
Mobility<-subset(data5, Mobility.Indexed == 1)
Cognitive<-subset(data5, Cognitive.Indexed == 1)
```

Next we look at the number of studies with each type of index
```{r}
Frailty %>%
  count(Study)

Mobility %>%
  count(Study)

Cognitive %>%
  count(Study)
```


## 5.1) Frailty

Social Costs
```{r}
#Societal Means
CI.intvl(Frailty$Societal_All)

#Societal SD Liberal 
CI.intvl(Frailty$Societal_All_lib_SD)
```

Healthcare Costs
```{r}
#Healthcare Means
CI.intvl(Frailty$Healthcare_All)

#Healthcare SD Liberal
CI.intvl(Frailty$Healthcare_All_lib_SD)
```

Personal Costs
```{r}
#Personal Means
CI.intvl(Frailty$Personal_All)

#Personal SD Liberal
CI.intvl(Frailty$Personal_All_lib_SD)
```


## 5.2) Mobility

Social Costs
```{r}
#Societal Means
CI.intvl(Mobility$Societal_All)

#Societal SD Liberal 
CI.intvl(Mobility$Societal_All_lib_SD)
```

Healthcare Costs
```{r}
#Healthcare Means
CI.intvl(Mobility$Healthcare_All)

#Healthcare SD Liberal
CI.intvl(Mobility$Healthcare_All_lib_SD)
```

Personal Costs
```{r}
#Personal Means
CI.intvl(Mobility$Personal_All)

#Personal SD Liberal
CI.intvl(Mobility$Personal_All_lib_SD)
```


## 5.3) Cognition

Social Costs
```{r}
#Societal Means
CI.intvl(Cognitive$Societal_All)

#Societal SD Liberal 
CI.intvl(Cognitive$Societal_All_lib_SD)
```

Healthcare Costs
```{r}
#Healthcare Means
CI.intvl(Cognitive$Healthcare_All)

#Healthcare SD Liberal
CI.intvl(Cognitive$Healthcare_All_lib_SD)
```

Personal Costs
```{r}
#Personal Means
CI.intvl(Cognitive$Personal_All)

#Personal SD Liberal
CI.intvl(Cognitive$Personal_All_lib_SD)
```